Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2019-04-30 doc/network.tex
Meeting with Daniel

The observed generation intervals are subject to right-censoring -- it is not possible to trace individuals that have not been incted yet.
During an ongoing outbreak, the observed generation-interval distribution will always have a shorter mean than the intrinsic-interval distribution due to right-censoring.
Near the end of an outbreak, the effect of right-censoring becomes negligible but the observed generation intervals are still shorter on average than the intrinsic generation intervals.
The remaining bias can be explained by depletion of the susceptible population.

@@

Susecptible depletion happening at all three levels shortens realized generation intervals but acts on different time scales.
Egocentric and local depletion effects are present from the beginning of an epidemic, even when depletion in the global susceptible population is negligible, and can strongly affect the initial spread of an epidemic.
Therefore, we predicted the observed generation intervals during an exponential growth phase to contain information about the contact structure, allowing us to estimate the \emph{effective} generation-interval distribution by simply accounting for the right-censoring.
Simulation studies confirmed our prediction: using the effective generation-interval distribution provides a proper link between $r$ and $\RR$.

We compare two methods for estimating the effective generation-interval distribution and assume that the effective generation-interval distribution follows a gamma distribution.
The gamma approximation of the generation interval distribution has widely used due to its simplicity;
we previously thought that summarizing the entire distribution with two moments (mean and variance) is sufficient to understand the role of generation-interval distributions in linking $r$ and $\RR$.
However, further investigation of our methods suggests that making a wrong distributional assumption leads to narrow confidence intervals, even though the gamma distribution looks indistinguishable from the shape of the intrinsic generation-interval distribution (derived from the SEIR model), making a wrong distributional assumption leads to narrow confidence intervals \swp{Appendix}.
These results are particularly alarming because it is impossible to know the true shape of the distribution [CITE].

@@

@@

>>>>

The observed generation intervals are subject to right-censoring -- it is not possible to trace individuals that have not been infected yet.
During an ongoing outbreak, the observed generation-interval distribution will always have a shorter mean than the intrinsic-interval distribution due to right-censoring.
Near the end of an outbreak, the effect of right-censoring becomes negligible but the observed generation intervals are still shorter on average than the intrinsic generation intervals.
The remaining bias can be explained by depletion of the susceptible population.

@@

Susceptible depletion happening at all three levels shortens realized generation intervals but acts on different time scales.
Egocentric and local depletion effects are present from the beginning of an epidemic, even when depletion in the global susceptible population is negligible, and can strongly affect the initial spread of an epidemic.
Therefore, we predicted the observed generation intervals during an exponential growth phase to contain information about the contact structure, allowing us to estimate the \emph{effective} generation-interval distribution by simply accounting for the right-censoring.
Simulation studies confirmed our prediction: using the effective generation-interval distribution provides the correct link between $r$ and $\RR$.

We compare two methods for estimating the effective generation-interval distribution and assume that the effective generation-interval distribution follows a gamma distribution.
The gamma approximation of the generation interval distribution has widely used due to its simplicity;
we previously thought that summarizing the entire distribution with two moments (mean and variance) is sufficient to understand the role of generation-interval distributions in linking $r$ and $\RR$.
However, further investigation of our methods suggests that making a wrong distributional assumption leads to overly narrow confidence intervals, even though the gamma distribution looks indistinguishable from the shape of the intrinsic generation-interval distribution (derived from the SEIR model) \swp{Appendix}.
These results are particularly alarming because it is impossible to know the true shape of the distribution [CITE].

@@

\jd{Let me try.}

@@

\jd{Maybe a bit more detail here.}
